---
title: "Prepare Forest Data with Helper Functions"
output:
  rmarkdown::html_vignette:
    highlight: pygments
vignette: >
  %\VignetteIndexEntry{Prepare Forest Data with Helper Functions}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 8,
  fig.height = 4.5
)
```

```{r setup}
library(ggforestplotR)
library(ggplot2)
```

This short article covers the `forest_data` interchange object and the helper
functions that prepare it before a plot is drawn.

## Use `as_forest_data()` to standardize a coefficient table

`as_forest_data()` converts your column names into the internal structure used
by `ggforestplotR`. The result is a `forest_data` data-frame subclass containing
the columns expected by `ggforestplot()`, `add_forest_table()`, and
`add_split_table()`.

```{r as-forest-data}
raw_coefs <- data.frame(
  variable = c("Age", "BMI", "Treatment"),
  beta = c(0.10, -0.08, 0.34),
  lower = c(0.02, -0.16, 0.12),
  upper = c(0.18, 0.00, 0.56),
  display = c("Age", "BMI", "Treatment"),
  section = c("Clinical", "Clinical", "Treatment"),
  sample_size = c(120, 115, 98),
  p_value = c(0.04, 0.15, 0.001)
)

forest_ready <- as_forest_data(
  data = raw_coefs,
  term = "variable",
  estimate = "beta",
  conf.low = "lower",
  conf.high = "upper",
  label = "display",
  grouping = "section",
  n = "sample_size",
  p.value = "p_value",
  estimate_scale = "identity",
  effect_label = "Beta",
  conf.level = 0.95
)
```

Use `forest_metadata()` to inspect the semantic and provenance contract. The
complete fitted model and a duplicate copy of the source data are not retained.

```{r inspect-forest-metadata}
forest_metadata(forest_ready)[c(
  "estimate_scale", "axis_transform", "effect_label", "conf_level",
  "reference_value", "source_model", "source_package", "source_columns"
)]
```

Once the data are standardized, you can pass them straight into `ggforestplot()`.

```{r helper-to-plot}
ggforestplot(forest_ready)
```

## Use `as_forest_data()` for model objects

If `broom` is available, the model-specific `as_forest_data()` method pulls
coefficient estimates and confidence limits from a fitted model. The method
also assigns model semantics such as `Beta`, `OR`, or `HR`.

```{r tidy-model}
fit <- lm(mpg ~ wt + hp + qsec, data = mtcars)

model_ready <- as_forest_data(fit)
```

The returned object can be passed directly into `ggforestplot()`.

```{r helper-to-plot-model}
ggforestplot(model_ready)
```

`tidy_forest_model()` remains available as a compatibility wrapper and returns
the same `forest_data` object.
