---
title: "Getting Started with ggforestplotR"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting Started with ggforestplotR}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 8,
  fig.height = 5
)
```

`ggforestplotR` is built to make forest plots feel like part of a normal
`ggplot2` workflow. You can start from a simple coefficient table, scale up to
grouped displays, attach a publication-friendly side table, and then keep
customizing the output with familiar `ggplot2` layers and themes.

This vignette walks through five increasingly capable patterns:

1. Build a basic forest plot from a coefficient table
2. Split rows into grouped sections
3. Attach a left or right summary table
4. Compare multiple estimates for the same terms
5. Add custom styling on top of the returned `ggplot` object

```{r setup}
library(ggforestplotR)
library(ggplot2)
```

## A basic forest plot

The simplest workflow is to start with a data frame that already contains a
term name, an estimate, and confidence limits.

```{r basic-data}
basic_coefs <- data.frame(
  term = c("Age", "BMI", "Treatment"),
  estimate = c(0.10, -0.08, 0.34),
  conf.low = c(0.02, -0.16, 0.12),
  conf.high = c(0.18, 0.00, 0.56)
)

basic_coefs
```

```{r basic-plot}
ggforestplot(
  basic_coefs,
  title = "Basic forest plot"
)
```

If your column names are different, map them explicitly with the `term`,
`estimate`, `conf.low`, and `conf.high` arguments.

```{r basic-remap}
renamed_coefs <- data.frame(
  variable = c("Age", "BMI", "Treatment"),
  beta = c(0.10, -0.08, 0.34),
  lower = c(0.02, -0.16, 0.12),
  upper = c(0.18, 0.00, 0.56)
)

ggforestplot(
  renamed_coefs,
  term = "variable",
  estimate = "beta",
  conf.low = "lower",
  conf.high = "upper",
  title = "Basic forest plot with remapped columns"
)
```

## Group rows into sections

Many forest plots read better when related variables are visually grouped
together. The `grouping` argument splits the plot into labeled sections.

```{r grouped-data}
sectioned_coefs <- data.frame(
  term = c("Age", "BMI", "Smoking", "Stage II", "Stage III", "Nodes"),
  estimate = c(0.10, -0.08, 0.20, 0.34, 0.52, 0.28),
  conf.low = c(0.02, -0.16, 0.05, 0.12, 0.20, 0.06),
  conf.high = c(0.18, 0.00, 0.35, 0.56, 0.84, 0.50),
  section = c(
    "Clinical", "Clinical", "Clinical",
    "Tumor", "Tumor", "Tumor"
  )
)

sectioned_coefs
```

```{r grouped-plot}
ggforestplot(
  sectioned_coefs,
  grouping = "section",
  title = "Forest plot with grouped sections"
)
```

This is also a good place to turn on alternating row striping.

```{r grouped-striped}
ggforestplot(
  sectioned_coefs,
  grouping = "section",
  striped_rows = TRUE,
  stripe_fill = "grey94",
  title = "Grouped sections with striped rows"
)
```

## Attach a side table

Many reporting workflows need both the visual forest plot and a compact text
summary beside it. `ggforestplot()` can attach a table to the left or right
side of the plot showing terms, optional N values, and a formatted estimate
column.

```{r table-data}
tabled_coefs <- data.frame(
  term = c("Age", "BMI", "Smoking", "Stage II", "Stage III"),
  estimate = c(0.12, -0.10, 0.18, 0.30, 0.46),
  conf.low = c(0.03, -0.18, 0.04, 0.10, 0.18),
  conf.high = c(0.21, -0.02, 0.32, 0.50, 0.74),
  sample_size = c(120, 115, 98, 87, 83),
  section = c("Clinical", "Clinical", "Clinical", "Tumor", "Tumor")
)
```

```{r table-left}
ggforestplot(
  tabled_coefs,
  grouping = "section",
  n = "sample_size",
  striped_rows = TRUE,
  table_position = "left",
  table_estimate_label = "Beta",
  title = "Forest plot with a left-side summary table"
)
```

You can also move the table to the right side and relabel the estimate header.

```{r table-right}
ggforestplot(
  tabled_coefs,
  grouping = "section",
  n = "sample_size",
  table_position = "right",
  table_estimate_label = "HR",
  title = "Forest plot with a right-side summary table"
)
```

When `table_position` is not `"none"`, the return value is a combined grob
instead of a plain `ggplot`, so the most flexible styling workflow is to finish
the plot styling first and then attach the table.

## Compare multiple estimates per term

The `group` argument is for a different kind of grouping: multiple estimates for
the same term. This is useful when comparing models, cohorts, or timepoints.

```{r comparison-data}
comparison_coefs <- data.frame(
  term = rep(c("Age", "BMI", "Smoking", "Stage II", "Stage III"), 2),
  estimate = c(
    0.12, -0.10, 0.18, 0.30, 0.46,
    0.08, -0.05, 0.24, 0.40, 0.58
  ),
  conf.low = c(
    0.03, -0.18, 0.04, 0.10, 0.18,
    0.00, -0.13, 0.10, 0.20, 0.30
  ),
  conf.high = c(
    0.21, -0.02, 0.32, 0.50, 0.74,
    0.16, 0.03, 0.38, 0.60, 0.86
  ),
  model = rep(c("Model A", "Model B"), each = 5),
  section = rep(c("Clinical", "Clinical", "Clinical", "Tumor", "Tumor"), 2),
  sample_size = rep(c(120, 115, 98, 87, 83), 2)
)

comparison_coefs
```

```{r comparison-plot}
ggforestplot(
  comparison_coefs,
  group = "model",
  grouping = "section",
  n = "sample_size",
  striped_rows = TRUE,
  dodge_width = 0.5,
  table_position = "right",
  table_estimate_label = "HR",
  title = "Comparing two model specifications"
)
```

When you pass a `group` column, `ggforestplot()` colors each series and dodges
the points so the estimates remain readable. In the attached table, grouped
rows are collapsed into a single text cell per term.

## Start from a fitted model

If you have a model object and `broom` is installed, `ggforestplot()` can call
`broom::tidy()` for you. This makes it easy to move from a fitted model to a
coefficient plot without building the table yourself first.

```{r model-plot}
fit <- lm(mpg ~ wt + hp + qsec, data = mtcars)

ggforestplot(
  fit,
  sort_terms = "descending",
  title = "Forest plot directly from an lm() object"
)
```

If you want to inspect or reuse the intermediate data, call
`tidy_forest_model()` directly.

```{r tidy-model}
tidy_forest_model(fit)
```

## Custom styling with ggplot2

The return value from `ggforestplot()` is a regular `ggplot` object when no side
table is attached, so you can continue styling it with scales, labels, and
themes.

```{r advanced-style}
ggforestplot(
  comparison_coefs,
  group = "model",
  grouping = "section",
  striped_rows = TRUE,
  stripe_fill = "#F3F4F6",
  zero_line_colour = "#9CA3AF",
  zero_line_linetype = 3,
  point_size = 2.8,
  line_size = 0.6,
  title = "Customized grouped forest plot",
  xlab = "Effect estimate",
  ylab = NULL
) +
  scale_colour_manual(
    values = c("Model A" = "#1D4ED8", "Model B" = "#D97706")
  ) +
  labs(
    subtitle = "Clinical and tumor terms shown in separate sections",
    caption = "Example data created for the vignette"
  ) +
  theme(
    legend.position = "top",
    panel.grid.major.y = element_blank(),
    strip.background = element_rect(fill = "#E5E7EB", colour = NA),
    strip.text.y.left = element_text(face = "bold"),
    plot.title.position = "plot"
  )
```

That last example is the main design idea behind `ggforestplotR`: the package
should handle the repetitive forest-plot mechanics, while still letting you
finish the visual design with the rest of the `ggplot2` ecosystem.